### 2024-04-04 paper/paper.qmd
Update paper and README

---
title: "The Impact of the 2022-23 NBA Rule Change"
subtitle: "Exploring the Relationship Between Player Shooting Performance, MVP Points, and the New Transitional Foul Rule"
author: 
  - Amie Liu
thanks: "Code and data are available at: https://github.com/AmieLiu/the-NBA-Rule-Change.git."
date: today
date-format: long
abstract: "Rule changes in professional basketball play a critical role in shaping both strategies and player performances. Specifically, the introduction of the transitional foul rule in the 2022-23 NBA season, which changes the way fouls are called during transition plays, could significantly affect players' shooting efficiency and their prospects for MVP awards. This paper examines the implications of this rule change, comparing player statistics between the 2022-23 and 2021-22 seasons. We find significant correlations between shooting efficiency and MVP points, highlighting the rule's influence on player recognition. This paper offers insights into the interplay between rule modifications, player performance, and individual accolades, underscoring the broader significance of adapting to regulatory shifts in professional sports."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
```

# Introduction

In professional basketball, rule changes can reshape the game's landscape, altering strategies, player performances, and ultimately, the trajectory of teams' successes. The National Basketball Association (NBA), as the premier basketball league globally, often implements rule modifications to refine gameplay, increase scoring opportunities, and foster a more open and free-flowing style of basketball [@nba]. Under the new transitional foul rule in the 2022-23 NBA season, take fouls committed during transition scoring opportunities will result in one free throw and continued possession for the offended team [@rule], encouraging defenders to make legitimate plays.

This paper analyzes the consequences of the 2022-23 NBA rule change, particularly its impact on player shooting performance and the dynamics of the Most Valuable Player (MVP) competition, comparing statistics with the previous season, 2021-22. Our estimand is to explore the relationship between player shooting performance and MVP points they won in these NBA seasons. We hypothesize that players who excel in transition offense may experience an increase in shooting efficiency as they capitalize on more fast-break opportunities without facing as many transition-take fouls, potentially leading to greater recognition in MVP voting due to the positive correlation between shooting efficiency and MVP consideration.

A clear gap exists in understanding the nuanced impacts of the new transitional foul rule on player performance and MVP considerations, especially when compared to the previous season. This paper seeks to fill this gap by thoroughly analyzing the statistical data from both the 2022-23 and 2021-22 NBA seasons to highlight the changes, continuities, and emerging trends over these times.

To address this gap, we used the R programming language [@citeR] to analyze shooting performance metrics, such as field goal percentage (FG%), three-point percentage (3P%), free throw percentage (FT%), as well as MVP voting points won from both seasons. Our findings reveal patterns in player shooting performance, MVP points accumulation, and the intertwined effects of the transitional foul rule… We discovered significant correlations between shooting efficiency and MVP points, with distinct variations observed between the two seasons.

Understanding the implications of the transitional foul rule in comparison to the previous season is crucial for stakeholders across the NBA landscape, from players and coaches to researchers and fans. By illuminating the intricate relationship between rule changes, player performance, and individual accolades, this paper contributes to a deeper understanding of the evolving dynamics within the league.

This paper is structured into three main sections: Data, Results, and Discussion. @sec-data Data explores sources and methodology. It includes various sources used for data collection, cleaning, and analyzing processes, drawing from the dataset. In @sec-results Results, we present our findings, focusing on data visualizations and analyses. @sec-discussion Discussion provides a comprehensive analysis of our research, highlights the implications for NBA basketball, the importance of limitations and considerations in our analysis, and identifies areas for future research. @sec-conclusion Conclusion summarizes our main findings and the broader significance of our research.

# Data {#sec-data}

Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-bills
#| fig-cap: Bills of penguins
#| echo: false

```

Talk more about it.

And also planes (@fig-planes). (You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data |> 
  ggplot(aes(x = width, y = length)) +
  geom_point(alpha = 0.8) +
  theme_minimal() +
  labs(x = "Wing width (mm)",
       y = "Wing length (mm)")
```

Talk way more about it.

# Model {#sec-model}

The goal of our modelling strategy is twofold. Firstly,...

Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results {#sec-results}

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

```

# Discussion {#sec-discussion}

## First discussion point {#sec-first-point}

If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.

## Second discussion point

## Third discussion point

## Weaknesses and next steps

Weaknesses and next steps should also be included.

\newpage

# Conclusion {#sec-conclusion}

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

\newpage

# References

### 2024-04-19 scripts/00-simulate_data.R
Update scripts, including 00 simulate data, 01 download data, and 02 data cleaning

@@ -1,19 +1,30 @@
 #### Preamble ####
-# Purpose: Simulates... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Simulates the dataset for 2021-22 and 2022-23 NBA Most Valuable Player 
+# Author: Amie Liu
+# Date: 19 April 2024
+# Contact: amie.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
 
 #### Simulate data ####
-# [...ADD CODE HERE...]
+# Set a random seed
+set.seed(123)
 
+simulated_data <-
+  tibble(
+    # Use 1 through to 10 to represent each rank
+    "Rank" = 1:10,
+    # Randomly pick an option, with replacement, 151 times
+    "Party" = sample(
+      x = c("Liberal", "Labor", "National", "Green", "Other"),
+      size = 151,
+      replace = TRUE
+    )
+  )
 
-
+simulated_data

### 2024-04-19 scripts/01-download_data.R

@@ -1,26 +1,20 @@
 #### Preamble ####
-# Purpose: Downloads and saves the data from [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Downloads and saves the data from Basketball Reference
+# Author: Amie Liu
+# Date: 19 April 2024
+# Contact: amie.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: None
 
 
-#### Workspace setup ####
-library(opendatatoronto)
-library(tidyverse)
-# [...UPDATE THIS...]
+#### Download and save data ####
+# Data is accessed from Basketball Reference website: https://www.basketball-reference.com/awards/awards_2023.html, https://www.basketball-reference.com/awards/awards_2022.html
+# Data needs to be downloaded manually
+# To download the data, I followed the instruction from Sports Reference Blog: https://www.sports-reference.com/blog/2016/11/exporting-data/
+# 1. Select "Get table as CSV (for Excel)" under the "Share & more" tab
+# 2. Copy and paste the entire table to Microsoft Excel
+# 3. Select "text to columns" function under the "Data" tab in Microsoft Excel
+# 4. Choose the delimiter "comma"
+# 5. Finish "text to columns" function
+# 6. Save the table as .csv file in the project data file
 
-#### Download data ####
-# [...ADD CODE HERE TO DOWNLOAD...]
-
-
-
-#### Save data ####
-# [...UPDATE THIS...]
-# change the_raw_data to whatever name you assigned when you downloaded it.
-write_csv(the_raw_data, "inputs/data/raw_data.csv") 
-
-         

### 2024-04-19 scripts/02-data_cleaning.R

@@ -1,44 +1,57 @@
 #### Preamble ####
-# Purpose: Cleans the raw plane data recorded by two observers..... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 6 April 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Cleans the 2021-22 and 2022-23 NBA Most Valuable Player datasets
+# Author: Amie Liu
+# Date: 19 April 2024
+# Contact: amie.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites: 
+#   01-download_data.R
+
 
 #### Workspace setup ####
 library(tidyverse)
+library(dplyr)
+library(janitor)
 
-#### Clean data ####
-raw_data <- read_csv("inputs/data/plane_data.csv")
+#### Read in data ####
+raw_data2023 <- read_csv("data/raw_data/2022-23_raw_data.csv")
+raw_data2022 <- read_csv("data/raw_data/2021-22_raw_data.csv")
 
-cleaned_data <-
-  raw_data |>
+
+#### Clean 2022-23 Data ####
+cleaned_data2023 <-
+  # select specific rows and columns
+  raw_data2023[2:11,c(1, 6, 16, 17, 18)] |>
   janitor::clean_names() |>
-  select(wing_width_mm, wing_length_mm, flying_time_sec_first_timer) |>
-  filter(wing_width_mm != "caw") |>
-  mutate(
-    flying_time_sec_first_timer = if_else(flying_time_sec_first_timer == "1,35",
-                                   "1.35",
-                                   flying_time_sec_first_timer)
+  # rename columns' names
+  rename(
+    rank = x1,
+    points_won = voting_6,
+    field_goal_percentage = shooting_16,
+    three_points_percentage = shooting_17,
+    free_throw_percentage = shooting_18
+  )
+
+
+#### Clean 2021-22 Data ####
+cleaned_data2022 <-
+  # select specific rows and columns
+  raw_data2022[2:11,c(1, 6, 16, 17, 18)] |>
+  janitor::clean_names() |>
+  # rename column names
+  rename(
+    rank = x1,
+    points_won = voting_6,
+    field_goal_percentage = shooting_16,
+    three_points_percentage = shooting_17,
+    free_throw_percentage = shooting_18
   ) |>
[…]

### 2024-04-19 scripts/02-data_cleaning.R
Update scripts, inclusing 02-data_cleaning and 03-test_data

@@ -13,7 +13,8 @@ library(tidyverse)
 library(dplyr)
 library(janitor)
 
-#### Read in data ####
+
+#### Read data ####
 raw_data2023 <- read_csv("data/raw_data/2022-23_raw_data.csv")
 raw_data2022 <- read_csv("data/raw_data/2021-22_raw_data.csv")
 
@@ -55,3 +56,4 @@ cleaned_data2022 <-
 #### Save data ####
 write_csv(cleaned_data2023, "data/analysis_data/analysis_data2023.csv")
 write_csv(cleaned_data2022, "data/analysis_data/analysis_data2022.csv")
+

### 2024-04-19 scripts/03-test_data.R

@@ -1,15 +1,56 @@
 #### Preamble ####
-# Purpose: Tests... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Tests data
+# Author: Amie Liu
+# Date: 19 April 2024
+# Contact: amie.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites:
+#   01-download_data.R
+#   02-data_cleaning.R
 
 
 #### Workspace setup ####
 library(tidyverse)
-# [...UPDATE THIS...]
+
+
+#### Read data ####
+cleaned_data2023 <- read_csv("data/analysis_data/analysis_data2023.csv")
+cleaned_data2022 <- read_csv("data/analysis_data/analysis_data2022.csv")
+
 
 #### Test data ####
+# Check if all column names match the specified names and order: "rank", "points_won", "field_goal_percentage", "three_points_percentage", "free_throw_percentage"
+all(colnames(cleaned_data2023) == c("rank", "points_won", "field_goal_percentage", "three_points_percentage", "free_throw_percentage"))
+all(colnames(cleaned_data2022) == c("rank", "points_won", "field_goal_percentage", "three_points_percentage", "free_throw_percentage"))
+
+# Check if the length of "rank" is 10
+cleaned_data2023$rank |> length() == 10
+cleaned_data2022$rank |> length() == 10
+
+# Check if the minimum value of "rank" is 1
+cleaned_data2023$rank |> min() == 1
+cleaned_data2022$rank |> min() == 1
+
+# Check if the maximum value of "rank" is 10
+cleaned_data2023$rank |> max() == 10
+cleaned_data2022$rank |> max() == 10
+
+# Check if the largest value of "points_won" is smaller or equal to 1000
+cleaned_data2023$points_won |> max() <= 1000
+cleaned_data2022$points_won |> max() <= 1000
+
+# Check if the smallest value of "points_won" is larger or equal to 0
+cleaned_data2023$points_won |> min() >= 0
+cleaned_data2022$points_won |> min() >= 0
+
+# Check if all values of "field_goal_percentage" is smaller than 1
[…]

### 2024-04-19 scripts/00-simulate_data.R
Update simulate_data

@@ -15,16 +15,18 @@ library(tidyverse)
 # Set a random seed
 set.seed(123)
 
-simulated_data <-
+simulated_data2023 <-
   tibble(
     # Use 1 through to 10 to represent each rank
     "Rank" = 1:10,
-    # Randomly pick an option, with replacement, 151 times
-    "Party" = sample(
-      x = c("Liberal", "Labor", "National", "Green", "Other"),
-      size = 151,
-      replace = TRUE
-    )
+    # Randomly pick a number between 0 and 1000
+    "Points Won" = sample(0:1000, 10, replace = TRUE),
+    # Randomly generate a number between 0 and 1 with three decimal places for field goal percentage 
+    "Field Goal Percentage" = round(runif(10, min = 0, max = 1) * 1000) / 1000,
+    # Randomly generate a number between 0 and 1 with three decimal places for three points percentage
+    "Three Points Percentage" = round(runif(10, min = 0, max = 1) * 1000) / 1000,
+    # Randomly generate a number between 0 and 1 with three decimal places for free throw percentage
+    "Free Throw Percentage" = round(runif(10, min = 0, max = 1) * 1000) / 1000
   )
 
-simulated_data
+simulated_data2023

### 2024-04-19 scripts/00-simulate_data.R
Update README and simulate_data

@@ -15,7 +15,8 @@ library(tidyverse)
 # Set a random seed
 set.seed(123)
 
-simulated_data2023 <-
+# The 2021-22 and 2022-23 seasons will have the same type of dataset, but the values will be different
+simulated_data <-
   tibble(
     # Use 1 through to 10 to represent each rank
     "Rank" = 1:10,
@@ -29,4 +30,4 @@ simulated_data2023 <-
     "Free Throw Percentage" = round(runif(10, min = 0, max = 1) * 1000) / 1000
   )
 
-simulated_data2023
+simulated_data

### 2024-04-20 paper/paper.qmd
Update table

@@
 title: "The Impact of the 2022-23 NBA Rule Change"
-subtitle: "Exploring the Relationship Between Player Shooting Performance, MVP Points, and the New Transitional Foul Rule"
+subtitle: "Exploring the Relationship Between the New Transitional Foul Rule, Player Shooting Performance, and MVP Points"
 author: 
@@
 library(tidyverse)
+library(knitr)
+
+# Read data
+analyzed_data2023 <- read_csv("../data/analysis_data/analysis_data2023.csv")
+analyzed_data2022 <- read_csv("../data/analysis_data/analysis_data2022.csv")
 ```
@@
 This paper seeks to fill this gap by thoroughly analyzing the statistical data from both the 2022-23 and 2021-22 NBA seasons to highlight the changes, continuities, and emerging trends over these times.
-
 To address this gap, we used the R programming language [@citeR] to analyze shooting performance metrics, such as field goal percentage (FG%), three-point percentage (3P%), free throw percentage (FT%), as well as MVP voting points won from both seasons.
@@
 
-Some of our data is of penguins (@fig-bills), from @palmerpenguins.
+Some of our data is of penguins (see @tbl-table2023 and @tbl-table2022).
 
 ```{r}
-#| label: fig-bills
-#| fig-cap: Bills of penguins
+#| label: tbl-table2023
+#| tbl-cap: 2022-2023 season most valuable player dataset
 #| echo: false
+#| warning: false
 
+# Display the "analyzed_data2023" dataset
+analyzed_data2023 |>
+  kable(
+    # Give each column a name
+    col.names = c("Rank", "MVP Points Won", "Field Goal Percentage", "Three Points Percentage", "Free Throw Percentage")
+  )
+```
+
+```{r}
+#| label: tbl-table2022
+#| tbl-cap: 2021-2022 season most valuable player dataset
+#| echo: false
+#| warning: false
+
+# Display the "analyzed_data2022" dataset
+analyzed_data2022 |>
+  kable(
+    # Give each column a name
+    col.names = c("Rank", "MVP Points Won", "Field Goal Percentage", "Three Points Percentage", "Free Throw Percentage")
+  )
 ```
@@
 
-analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
+#analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
 
-analysis_data |> 
-  ggplot(aes(x = width, y = length)) +
-  geom_point(alpha = 0.8) +
-  theme_minimal() +
-  labs(x = "Wing width (mm)",
-       y = "Wing length (mm)")
+#analysis_data |> 
+#  ggplot(aes(x = width, y = length)) +
+#  geom_point(alpha = 0.8) +
+#  theme_minimal() +
+#  labs(x = "Wing width (mm)",
+#       y = "Wing length (mm)")
 ```
@@
 \appendix
-
 # Appendix {-}

### 2024-04-20 scripts/00-simulate_data.R

@@ -20,8 +20,8 @@ simulated_data <-
   tibble(
     # Use 1 through to 10 to represent each rank
     "Rank" = 1:10,
-    # Randomly pick a number between 0 and 1000
-    "Points Won" = sample(0:1000, 10, replace = TRUE),
+    # Randomly pick a number between 0 and 1000 for MVP points won
+    "MVP Points Won" = sample(0:1000, 10, replace = TRUE),
     # Randomly generate a number between 0 and 1 with three decimal places for field goal percentage 
     "Field Goal Percentage" = round(runif(10, min = 0, max = 1) * 1000) / 1000,
     # Randomly generate a number between 0 and 1 with three decimal places for three points percentage

### 2024-04-20 paper/paper.qmd
Update bar and scatter plots

@@
 library(knitr)
+library(dplyr)
+library(cowplot)
 
@@
 #| label: tbl-table2023
-#| tbl-cap: 2022-2023 season most valuable player dataset
+#| tbl-cap: 2022-2023 season most valuable player table 
 #| echo: false
@@
 #| label: tbl-table2022
-#| tbl-cap: 2021-2022 season most valuable player dataset
+#| tbl-cap: 2021-2022 season most valuable player table
 #| echo: false
@@
 
-Our results are summarized in @tbl-modelresults.
+Our results are summarized in @tbl-modelresults. (see @fig-points)
 
@@
 #| warning: false
+
+```
+
+```{r}
+#| label: fig-points
+#| fig-cap: Number of MVP points won by each player from the top 10 rankings in the 2021-22 and 2022-23 seasons
+#| fig.width: 8 
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create a bar chart to visualize the distribution of MVP points won by rank based on the analyzed_data2023 and analyzed_data2022 datasets
+
+# Add a "Season" column to the 2022-23 dataset
+analyzed_data2023 <- analyzed_data2023 |>
+  mutate(season = "2022-23")
+
+# Add a "Season" column to the 2021-22 dataset
+analyzed_data2022 <- analyzed_data2022 |>
+  mutate(season = "2021-22")
+
+# Combine the 2022-23 and 2021-22 datasets
+combined_data <- bind_rows(analyzed_data2023, analyzed_data2022)
+
+# Create the bar plot with rank on the x-axis and season as fill color
+combined_data |>
+  ggplot(aes(x = factor(rank), fill = factor(season))) +
+  # Add a grouped bar layer to create the grouped bar chart
+  geom_bar(position = "dodge", aes(y = points_won), stat = "identity") +
+  # Add text labels for the number of points won on top of each bar
+  geom_text(aes(label = points_won, y = points_won), vjust = -0.5, size = 3, position = position_dodge(width = 0.9)) +
+  # Scale the x-axis labels to display "Rank"
+  scale_x_discrete("Rank") +
+  # Scale the fill color to display "season" and specify the break points: 2021-22 Season", "2022-23 Season"
+  scale_fill_discrete(name = "Season",
+                      breaks = c("2021-22", "2022-23"),
+                      labels = c("2021-22 Season", "2022-23 Season")) +
+  # Apply a minimal theme to the chart
+  theme_minimal() +
+  # Label y-axis as "Number of MVP Points Won"
+  labs(y = "Number of MVP Points Won")
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create the scatter plot with trend line or smoothed line for "field_goal_percentage"
+scatterplot1 <- 
+  ggplot(combined_data, aes(x = rank, y = field_goal_percentage, color = season)) +
+  geom_point() +  # Add points for each data point
+  geom_smooth(method = "lm", se = FALSE) +  # Add a trend line (linear regression)
+  # Scale the x-axis labels to display "Rank"
+  scale_x_continuous(breaks = 1:10, labels = 1:10) +
+  # Scale the y-axis labels to display "Field Goal Percentage"
+  scale_y_continuous("Field Goal Percentage") +
+  # Change legend title, position, and x-axis
+  labs(color = "Season", x = "Rank") +
+  theme_minimal()
+
+```
+
+```{r}
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create the scatter plot with trend line or smoothed line for "three_points_percentage"
+scatterplot2 <- 
+  ggplot(combined_data, aes(x = rank, y = three_points_percentage, color = season)) +
+  geom_point() +  # Add points for each data point
+  geom_smooth(method = "lm", se = FALSE) +  # Add a trend line (linear regression)
+  # Scale the x-axis labels to display "Rank"
+  scale_x_continuous(breaks = 1:10, labels = 1:10) +
+  # Scale the y-axis labels to display "Three Points Percentage"
+  scale_y_continuous("Three Points Percentage") +
+  # Change legend title,  position, and x-axis
+  labs(color = "Season", x = "Rank") +
+  theme_minimal()
+
+```
+
+see (@fig-percentage)
+
+```{r}
+#| label: fig-percentage
+#| fig-cap: "Different shooting percentage by each player from the top 10 rankings in the 2021-22 and 2022-23 seasons"
+#| fig.height: 10
+#| fig.width: 8 
+#| echo: false
+#| warning: false
+#| message: false
+
+# Create the scatter plot with trend line or smoothed line for "free_throw_percentage"
+scatterplot3 <-
+  ggplot(combined_data, aes(x = rank, y = free_throw_percentage, color = season)) +
+  geom_point() +  # Add points for each data point
+  geom_smooth(method = "lm", se = FALSE) +  # Add a trend line (linear regression)
+  # Scale the x-axis labels to display "Rank"
+  scale_x_continuous(breaks = 1:10, labels = 1:10) +
+  # Scale the y-axis labels to display "Free Throw Percentage"
+  scale_y_continuous("Free Throw Percentage") +
+  # Change legend title, position, and x-axis
+  labs(color = "Season", x = "Rank") +
+  theme_minimal()
+
+# Create the combined plot
+combined_plot <- plot_grid(scatterplot1, scatterplot2, scatterplot3, ncol = 1)
+
+combined_plot
 

### 2024-04-21 paper/paper.qmd
Update paper, README, models

@@
 
-```{r}
-#| label: fig-planes
-#| fig-cap: Relationship between wing length and width
-#| echo: false
-#| warning: false
-#| message: false
-
-#analysis_data <- read_csv(here::here("data/analysis_data/analysis_data.csv"))
-
-#analysis_data |> 
-#  ggplot(aes(x = width, y = length)) +
-#  geom_point(alpha = 0.8) +
-#  theme_minimal() +
-#  labs(x = "Wing width (mm)",
-#       y = "Wing length (mm)")
-```
-
 Talk way more about it.
 
-# Model {#sec-model}
-
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
-
-## Model set-up
-
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
-
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
-
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
-
-### Model justification
-
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
-
-We can use maths by including latex between dollar signs, for instance $\theta$.
-
-# Results {#sec-results}
-
-Our results are summarized in @tbl-modelresults. (see @fig-points)
-
-```{r}
-#| echo: false
-#| eval: true
-#| warning: false
-#| message: false
-
-first_model <-
-  readRDS(file = here::here("models/first_model.rds"))
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
-#| warning: false
-
-```
-
+## Data Visualization
+### The distribution of MVP points won by top 10 players in the 2021-22 and 2022-23 seasons
 ```{r}
@@
 
+### Different shooting performances by top 10 players in the 2021-22 and 2022-23 seasons
 ```{r}
@@
 
-see (@fig-percentage)
-
 ```{r}
@@
 
+```
+
+# Model {#sec-model}
+
+The goal of our modelling strategy is twofold. Firstly,...
+
+Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+
+## Model set-up
+
+Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+
+\begin{align} 
+y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
+\mu_i &= \alpha + \beta_i + \gamma_i\\
+\alpha &\sim \mbox{Normal}(0, 2.5) \\
+\beta &\sim \mbox{Normal}(0, 2.5) \\
+\gamma &\sim \mbox{Normal}(0, 2.5) \\
+\sigma &\sim \mbox{Exponential}(1)
+\end{align}
+
+We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+
+### Model justification
+
+We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+
+We can use maths by including latex between dollar signs, for instance $\theta$.
+
+# Results {#sec-results}
+
+Our results are summarized in @tbl-modelresults. (see @fig-points)
+
+```{r}
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+model2023 <-
+  readRDS(file = "../models/model2023.rds")
+
+```
+
+```{r}
+#| echo: false
+#| eval: true
+#| label: tbl-modelresults
+#| tbl-cap: "Summary of the model results"
+#| warning: false
+
+# Create a data frame containing the model summary information
+model_summary <- tibble(
+  "Variable" = c("(Intercept)", "field_goal_percentage", "three_points_percentage", "free_throw_percentage"),
+  "Estimate" = c(-4628, 7399, 7410, -1640),
+  "Std. Error" = c(2004, 1994, 3740, 1417),
+  "t value" = c(-2.310, 3.712, 1.981, -1.158),
+  "Pr(>|t|)" = c(0.06028, 0.00995, 0.09485, 0.29093)
+)
+
+# Convert the data frame to a modelsummary table
+summary_table <- modelsummary(model_summary)
+
+summary_table
+
+```
+
+```{r}
+#| label: fig-modelresults
+#| fig-cap: "Observed vs. predicted values of MVP points won"
+#| fig.height: 10
+#| fig.width: 8 
+#| echo: false
+#| warning: false
+#| message: false
+
+# Diagnostic plots
+par(mfrow = c(2, 2))  # Arrange plots in a 2x2 grid
+plot(model2023)  # Generate plots
+
+# Create a data frame of observed vs. predicted values
+predicted_values <- data.frame(Observed = analyzed_data2023$points_won, 
+                               Predicted = predict(model2023))
+
+# Plot observed vs. predicted values
+ggplot(predicted_values, aes(x = Observed, y = Predicted)) +
+  geom_point() +
+  geom_abline(color = "red", linetype = "dashed") +  # Line x=y for reference
+  labs(title = "Observed vs. Predicted MVP Voting Points",
+       x = "Observed Points",
+       y = "Predicted Points") +
+  theme_minimal()
 ```

### 2024-04-21 scripts/04-model.R

@@ -1,11 +1,12 @@
 #### Preamble ####
-# Purpose: Models... [...UPDATE THIS...]
-# Author: Rohan Alexander [...UPDATE THIS...]
-# Date: 11 February 2023 [...UPDATE THIS...]
-# Contact: rohan.alexander@utoronto.ca [...UPDATE THIS...]
+# Purpose: Models data to predict total MVP voting points based on shooting performance (Field Goal Percentage, Three-Point Percentage, and Free Throw Percentage)
+# Author: Amie Liu
+# Date: 21 April 2024
+# Contact: amie.liu@mail.utoronto.ca
 # License: MIT
-# Pre-requisites: [...UPDATE THIS...]
-# Any other information needed? [...UPDATE THIS...]
+# Pre-requisites:
+#   01-download_data.R
+#   02-data_cleaning.R
 
 
 #### Workspace setup ####
@@ -13,25 +14,20 @@ library(tidyverse)
 library(rstanarm)
 
 #### Read data ####
-analysis_data <- read_csv("data/analysis_data/analysis_data.csv")
+analyzed_data2023 <- read_csv("data/analysis_data/analysis_data2023.csv")
+
 
 ### Model data ####
-first_model <-
-  stan_glm(
-    formula = flying_time ~ length + width,
-    data = analysis_data,
-    family = gaussian(),
-    prior = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_intercept = normal(location = 0, scale = 2.5, autoscale = TRUE),
-    prior_aux = exponential(rate = 1, autoscale = TRUE),
-    seed = 853
-  )
+# Fit a linear regression model to predict 'points_won' using FG%, 3P%, and FT%
+model2023 <- lm(points_won ~ field_goal_percentage + three_points_percentage + free_throw_percentage, data = analyzed_data2023)
+
+# Check the summary of the model to understand the fit and see the coefficients
+summary(model2023)
 
 
-#### Save model ####
+### Save model ####
 saveRDS(
-  first_model,
-  file = "models/first_model.rds"
+  model2023,
+  file = "models/model2023.rds"
 )
-
 

### 2024-04-21 paper/paper.qmd
Update model

@@
 library(cowplot)
+library(modelsummary)
+library(kableExtra)
 
@@
 
-# Create a data frame containing the model summary information
-model_summary <- tibble(
-  "Variable" = c("(Intercept)", "field_goal_percentage", "three_points_percentage", "free_throw_percentage"),
-  "Estimate" = c(-4628, 7399, 7410, -1640),
-  "Std. Error" = c(2004, 1994, 3740, 1417),
-  "t value" = c(-2.310, 3.712, 1.981, -1.158),
-  "Pr(>|t|)" = c(0.06028, 0.00995, 0.09485, 0.29093)
+# Generate a summary table with numerical values formatted to 2 decimal places
+model_table <- modelsummary(
+  list("2022-23 Season" = model2023),
+  fmt = 2
 )
 
-# Convert the data frame to a modelsummary table
-summary_table <- modelsummary(model_summary)
-
-summary_table
-
+model_table
 ```

### 2024-04-21 paper/paper.qmd
Update paper

@@
 library(cowplot)
-library(modelsummary)
-library(kableExtra)
 
@@
   labs(y = "Number of MVP Points Won")
-
 ```
@@
   theme_minimal()
-
 ```
@@
 combined_plot
-
 ```
@@
 
-Our results are summarized in @tbl-modelresults. (see @fig-points)
+Our results are summarized in @tbl-modelresults. (see @fig-points and @fig-modelresults)
 
@@
   readRDS(file = "../models/model2023.rds")
-
-```
-
-```{r}
-#| echo: false
-#| eval: true
-#| label: tbl-modelresults
-#| tbl-cap: "Summary of the model results"
-#| warning: false
-
-# Generate a summary table with numerical values formatted to 2 decimal places
-model_table <- modelsummary(
-  list("2022-23 Season" = model2023),
-  fmt = 2
-)
-
-model_table
 ```
@@
 # Diagnostic plots
-par(mfrow = c(2, 2))  # Arrange plots in a 2x2 grid
-plot(model2023)  # Generate plots
+#par(mfrow = c(2, 2))  # Arrange plots in a 2x2 grid
+#plot(model2023)  # Generate plots
 
@@
 # Plot observed vs. predicted values
+observed_predicted_values <-
   ggplot(predicted_values, aes(x = Observed, y = Predicted)) +
@@
   theme_minimal()
+
+observed_predicted_values
 ```

### 2024-04-21 paper/paper.qmd
Update tables

@@
 library(cowplot)
+library(broom)
 
@@
   # Change legend title, position, and x-axis
-  labs(color = "Season", x = "Rank") +
+  labs(title = "Field Goal Percentage by Rank", color = "Season", x = "Rank") +
   theme_minimal()
@@
   # Change legend title,  position, and x-axis
-  labs(color = "Season", x = "Rank") +
+  labs(title = "Three Points Percentage by Rank", color = "Season", x = "Rank") +
   theme_minimal()
@@
   # Change legend title, position, and x-axis
-  labs(color = "Season", x = "Rank") +
+  labs(title = "Free Throw Percentage by Rank", color = "Season", x = "Rank") +
   theme_minimal()
@@
 ```{r}
+#| label: tbl-coefficients
+#| tbl-cap: "Summary of regression model coefficients"
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Generate a tidy summary of the model
+model_summary <- tidy(model2023) |>
+  rename(
+    Term = term,
+    Estimate = estimate,
+    `Std. Error` = std.error,
+    Statistic = statistic,
+    `P-value` = p.value,
+  ) |>
+  # Rename the "Term" values
+  mutate(
+    Term =
+      case_match(
+        Term,
+        "(Intercept)" ~ "(Intercept)",
+        "field_goal_percentage" ~ "Field Goal Percentage",
+        "three_points_percentage" ~ "Three Points Percentage",
+        "free_throw_percentage" ~ "Free Throw Percentage"
+        )
+    )
+
+# Use knitr's kable function to format tables
+kable(model_summary)
+```
+
+```{r}
+#| label: tbl-additional
+#| tbl-cap: "Summary of additional regression model metrics"
+#| echo: false
+#| eval: true
+#| warning: false
+#| message: false
+
+# Include additional model metrics
+additional_metrics <- data.frame(
+  Metric = c("Residual Standard Error", "Multiple R-squared", "Adjusted R-squared", "F-statistic", "p-value"),
+  Value = c("314.3", "0.4446", "0.1669", "1.601", "0.2852")
+)
+
+kable(additional_metrics)
+```
+
+```{r}
 #| label: fig-modelresults
 #| fig-cap: "Observed vs. predicted values of MVP points won"
-#| fig.height: 10
 #| fig.width: 8 
@@
   geom_abline(color = "red", linetype = "dashed") +  # Line x=y for reference
-  labs(title = "Observed vs. Predicted MVP Voting Points",
-       x = "Observed Points",
+  labs(x = "Observed Points",
        y = "Predicted Points") +

### 2024-04-22 paper/paper.qmd
Update model section of paper, llm usage, references

@@
 
-The goal of our modelling strategy is twofold. Firstly,...
-
-Here we briefly describe the Bayesian analysis model used to investigate... Background details and diagnostics are included in [Appendix -@sec-model-details].
+To explore the impact of the 2022-23 NBA transitional foul rule on player shooting performance and its subsequent effect on MVP voting, we constructed a linear regression model.
+The objective is to quantify the relationship between shooting performance metrics and MVP voting outcomes.
 
@@
 
-Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.
+The dependent variable is MVP Voting Points and the independent variables are Field Goal Percentage (FG%), Three-Point Field Goal Percentage (3P%), and Free Throw Percentage (FT%).
+Each independent variable was chosen based on its potential influence on a player's performance under the conditions created by the new foul rule.
+We run the model in R [@citeR] utilizing the `rstanarm` package [@rstanarm].
+Our model is constructed as follows:
 
-\begin{align} 
-y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
-\mu_i &= \alpha + \beta_i + \gamma_i\\
-\alpha &\sim \mbox{Normal}(0, 2.5) \\
-\beta &\sim \mbox{Normal}(0, 2.5) \\
-\gamma &\sim \mbox{Normal}(0, 2.5) \\
-\sigma &\sim \mbox{Exponential}(1)
-\end{align}
+$$
+Y = \beta_0 + \beta_1X_1 + \beta_2X_2 + \beta_3X_3 + \epsilon
+$$
 
-We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.
+where:
+
+- $Y$ is the MVP Voting Points. It is the dependent variable that represents the points a player receives in the MVP voting process.
+- $\beta_0$ is the intercept of the regression model that represents the expected value of Total MVP Voting Points when all predictor variables (FG%, 3P%, FT%) are at zero.
+- $\beta_1$ is the coefficient for FG%. It indicates how changes in a player’s overall shooting efficiency from the field affect their MVP voting points.
+- $\beta_2$ is the the coefficient for 3P%. This reflects the impact of three-point shooting efficiency on MVP voting.
+- $\beta_3$ is the coefficient for FT%. This measures the influence of free throw shooting efficiency on MVP voting points.
+- $X_1$ measures the FG% made out of attempts.
+- $X_2$ measures the 3P% made out of attempts.
+- $X_3$ measures the FT% made out of attempts.
+- $\epsilon$ is the error term, representing the residual element not explained by the model. This captures the variability in MVP voting points not explained by shooting percentages alone.
 
@@
 
-We expect a positive relationship between the size of the wings and time spent aloft. In particular...
+The selection of linear regression for our analysis suggests linear relationships between the shooting metrics and MVP voting points.
+We hypothesized that improvements in these shooting metrics, reflecting enhanced performance in transition plays free from take fouls, would correlate positively with higher MVP voting scores.
+The use of FG%, 3P%, and FT% as predictors is justified by their direct impact on game performance, which is critical in MVP considerations.
+Moreover, these metrics are commonly used in basketball analytics to assess player efficiency and effectiveness, making them relevant for evaluating the consequences of rule changes.
 
-We can use maths by including latex between dollar signs, for instance $\theta$.
+Statistical significance of the model coefficients will be tested using t-tests, where a p-value less than 0.05 will indicate significant influence of the predictor on MVP voting points.
+The overall fit of the model will be assessed by the R-squared value, which explains the proportion of variance in MVP points accounted for by the predictors.
 

### 2024-04-22 paper/paper.qmd
Update paper, llm usage, and references
Delete 05 replication script

@@
 
-Some of our data is of penguins (see @tbl-table2023 and @tbl-table2022).
+The data for this paper are obtained from Basketball Reference [@basketball], which provides a wide range of basketball-related data that includes statistics, player and team information, game results, and more.
+The datasets used here are @basketball2023 and @basketball2022.
+These datasets include various statistics and information about NBA players who received votes for the MVP award, as well as their performance metrics for the 2022-23 and 2021-22 seasons.
+The data analysis involved using the R programming language [@citeR], along with the `tidyverse` [@citeTidyverse], `janitor` [@citeJanitor], `knitr` [@citeKnitr], `cowplot` [@cowplot], `broom` [@broom], and `rstanarm` [@rstanarm] packages for data collection, cleaning, analysis, and statistical modeling.
+Within the `tidyverse` package [@citeTidyverse], I used `readr` [@citeReadr], `dplyr` [@citeDplyr], `tidyr` [@citeTidyr], and `ggplot2` [@citeGgplot2] for reading CSV files, manipulating data, tidying data, and creating figures.
+
+## NBA 2021-22 and 2022-23 Seasons Top 10 MVP Data
+
+@tbl-table2023 presents the performance metrics of NBA players who received votes for the Most Valuable Player (MVP) award in the 2022-23 season.
+It includes a total of 10 rows and five variables: ranking of the top 10 players (Rank), the total number of MVP points won by each player (MVP Points Won), FG% (Field Goal Percentage), TP% (Three-Point Percentage), and FT% (Free Throw Percentage).
 
@@
 #| label: tbl-table2023
-#| tbl-cap: 2022-2023 season most valuable player table 
+#| tbl-cap: 2022-2023 season top 10 MVP table 
 #| echo: false
@@
 
+Field Goal Percentage indicates the efficiency of a player's shooting from the field, calculated as the ratio of successful field goals made to total field goal attempts.
+A higher field goal percentage suggests better shooting accuracy.
+Moreover, Three Points Percentage indicates the accuracy of a player's three-point shooting, measured as the percentage of successful three-point shots made out of total attempts.
+A higher three-point percentage reflects proficiency in long-range shooting.
+In addition, Free Throw Percentage represents the proficiency of a player in making free throws, calculated as the ratio of successful free throws made to total free throw attempts.
+A higher free throw percentage indicates better accuracy and consistency in free throw shooting.
+
+@tbl-table2022 presents performance metrics of NBA players who received votes for the MVP award during the 2021-22 season.
+This dataset mirrors the structure of the previously mentioned dataset (see @tbl-table2023), providing insights into key performance indicators for players in the 2021-22 season.
+
 ```{r}
 #| label: tbl-table2022
-#| tbl-cap: 2021-2022 season most valuable player table
+#| tbl-cap: 2021-2022 season top 10 MVP table
 #| echo: false
@@
 
-Talk more about it.
+## Data Measurement
 
-And also planes (@fig-planes).
-(You can change the height and width, but don't worry about doing that until you have finished every other aspect of the paper - Quarto will try to make it look nice and the defaults usually work well once you have enough text.)
-
-Talk way more about it.
+The measurement process quantifies real-world phenomena related to NBA player performance and translates them into dataset entries.
+We focused on performance metrics during the 2021-22 and 2022-23 seasons, including field goal percentage, three-point percentage, and free throw percentage.
+Data were collected from reputable sources like @basketball.
+These raw data were then processed and cleaned to ensure accuracy and consistency.
+The selected variables were defined and operationalized.
+For example, field goal percentage was calculated as the ratio of successful field goals made to attempts.
+Similarly, three-point and free throw percentages were computed using standardized formulas.
+These measurements became dataset entries, with rows representing players and columns representing performance metrics.
 
@@
 ### The distribution of MVP points won by top 10 players in the 2021-22 and 2022-23 seasons
+
+@fig-points illustrates the MVP points won by the top 10 NBA players in each of the 2022-23 and 2021-22 seasons.
+It clearly shows that the highest number of MVP points, which totaled 915, was given to the player who ranked 1st in the 2022-23 season.
+In contrast, the minimum point (1 point) was given to the player who ranked 10th in the 2021-22 season.
+
+When examining the MVP points across the corresponding ranks in both seasons, players in seven out of the ten ranks during the 2022-23 season received more points than their counterparts in the same ranks during the 2021-22 season.
+The top ranks in the 2022-23 season (Rank 1, 3, 4) generally received more points than in the previous season.
+This might indicate a shift in voting behavior or a stronger consensus among voters regarding the leading candidates.
+Players at lower ranks (6-10) showed a relatively small variation in points, suggesting that while the top contenders might shift significantly in performance or perception, those at lower ranks maintain a more consistent level of recognition.
+
 ```{r}
@@
 ### Different shooting performances by top 10 players in the 2021-22 and 2022-23 seasons
+
+@fig-percentage provides a comparison of the shooting performances of the top 10 NBA players across 2022-23 and 2021-22 seasons, including Field Goal, Three-Point, and Free Throw percentages.
+Each metric is depicted with trend lines to illustrate the general trends across player ranks from 1 to 10.
+
+The highest Field Goal Percentage observed was 0.632 in the 2022-23 season.
+In contrast, the lowest percentage was recorded at 0.437 in the 2021-22 season.
+The trend lines show that the 2022-23 season has higher percentages across the ranks compared to the previous season, despite a slight decline in percentage from the highest to the lowest ranked players in both seasons.
+This means there is an enhancement in field goal accuracy from the 2021-22 season to the 2022-23 season.
+
+For the Three-Point Percentage, the highest was recorded at 0.427 in the 2022-23 season, while the lowest was also observed in the same season at 0.275.
+Interestingly, the trend line for 2022-23 shows a slight increase as player rank increases, unlike in 2021-22, where the trend line remains relatively flat across ranks.
+There is a marginal increase in three-point shooting percentages in the 2022-23 season
+
+The scatter plot for Free Throw Percentage shows the highest performance at 0.923 in the 2021-22 season and the lowest at 0.645 in the 2022-23 season.
+Both seasons exhibit a slight upward trend in free throw efficiency from rank 1 to 10.
+The free throw percentage remains nearly constant in both seasons, with a minimal increase in the 2021-22 season.
+
 ```{r}
@@
 
-Our results are summarized in @tbl-modelresults. (see @fig-points and @fig-modelresults)
+Our results are summarized in @tbl-coefficients and @tbl-additional (see @fig-modelresults)
 

### 2024-04-23 paper/paper.qmd
Update model and results section
Add data analysis

@@
 - $Y$ is the MVP Voting Points. It is the dependent variable that represents the points a player receives in the MVP voting process.
+- $X_1$ measures the FG% made out of attempts.
+- $X_2$ measures the 3P% made out of attempts.
+- $X_3$ measures the FT% made out of attempts.
 - $\beta_0$ is the intercept of the regression model that represents the expected value of Total MVP Voting Points when all predictor variables (FG%, 3P%, FT%) are at zero.
@@
 - $\beta_3$ is the coefficient for FT%. This measures the influence of free throw shooting efficiency on MVP voting points.
-- $X_1$ measures the FG% made out of attempts.
-- $X_2$ measures the 3P% made out of attempts.
-- $X_3$ measures the FT% made out of attempts.
 - $\epsilon$ is the error term, representing the residual element not explained by the model. This captures the variability in MVP voting points not explained by shooting percentages alone.
@@
 # Results {#sec-results}
+## Summary of regression model results
+Our results are summarized in @tbl-coefficients and @tbl-additional.
+The estimate intercept is -920.99 with a high standard error of 1833.86 and high p-value of 0.63 (significantly greater than 0.05).
+The negative intercept might initially suggest a baseline negative effect, but considering high standard error and p-value, the intercept is not statistically significant.
+This means that the intercept does not provide reliable information about the baseline level of MVP points when shooting percentages are not accounted for.
 
-Our results are summarized in @tbl-coefficients and @tbl-additional (see @fig-modelresults)
+Field Goal Percentage (FG%) has a positive coefficient, indicating that an increase in FG% will have an increase in MVP points.
+The estimate suggests that for each percentage point increase in FG%, MVP points increase by approximately 3432.88.
+However, the standard error of 2139, and the p-value of 0.160, which is above the significant level of 0.05, suggests that this relationship is not statistically significant at the conventional level.
+Thus, while FG% shows a positive impact on MVP points, this effect is not strong enough to be considered reliable within this model.
+
+The coefficient for Three Points Percentage (3P%) is negative (-6037.23), suggesting that a higher 3P% correlates with fewer MVP points.
+It indicates a substantial decrease in MVP points for each percentage increase in 3P%.
+However, similar to FG%, the p-value of 0.140 indicates that this negative relationship is not statistically significant.
+The large standard error further points to high variability in the data, leading to uncertainty in the estimate.
+
+For Free Throw Percentage (FT%), the regression model estimates a coefficient of 1828, which shows a positive relationship between FT% and MVP points.
+This suggests that higher FT% could be associated with more MVP points.
+However, the p-value of 0.356 indicates that this effect is not statistically significant, which exceeds the commonly accepted 0.05 for declaring statistical significance.
+Moreover, the standard error (1828.51) is nearly equal to the estimate (1827.86), indicating low precision in the estimate.
+Consequently, there is considerable uncertainty around its effect on MVP points.
 
@@
     Statistic = statistic,
-    `P-value` = p.value,
+    `p-value` = p.value,
   ) |>
+  # Format numerical columns to the second decimal place
+  mutate(across(where(is.numeric), ~format(round(., 2), nsmall = 2))) |>
   # Rename the "Term" values
@@
 
+@tbl-additional provides several important statistical metrics that help evaluate the overall fit and effectiveness of the regression model.
+The high Residual Standard Error of 314.30 suggests substantial unexplained variability, indicating the model's predictions can deviate significantly from actual values.
+While the Multiple R-squared value of 0.44 shows that the model explains about 44% of the variability in MVP points, the lower Adjusted R-squared of 0.17 indicates that the model's explanatory power significantly reduces when adjusted for the number of predictors, suggesting potential overfitting or irrelevance of some variables.
+Furthermore, the F-statistic of 1.60, which is relatively low, and p-value of 0.29 indicate that the model is not statistically significant, implying that the predictors do not reliably explain variations in MVP points.
+
 ```{r}
@@
 
+# Round the Value column to the second decimal place
+additional_metrics$Value <- round(as.numeric(additional_metrics$Value), 2)
+
 kable(additional_metrics)
 ```
+
+## Observed versus predicted MVP points won
+@fig-modelresults shows the observed MVP points (actual outcomes) against predicted MVP points (outcomes estimated by the model).
+The perfect predictions would result in all points falling exactly on the red dashed line, where the predicted values equal the observed values.
+However, as we can see, the data points are notably distant from this line, particularly at higher observed points.
+For instance, an observed value of 915 points in a predicted value of only 534.42.
+This significant deviation indicates that the model tends to underpredict MVP points for players with higher observed scores.
+In addition, for lower observed points, the model's performance is inconsistent.
+For instance, when the observed value is at 30, the predicted value is slightly negative (-5.09), which doesn't align with the possible range of MVP points.
+Conversely, an observed value of 3 has an inaccurate predicted value of 369.98.
+This means there is a potential issue of reliability and precision in the model’s predictions for top-performing players.
 
@@
 #| message: false
-
-# Diagnostic plots
-#par(mfrow = c(2, 2))  # Arrange plots in a 2x2 grid
-#plot(model2023)  # Generate plots
 
@@
 
-\appendix
-# Appendix {-}
-
-# Additional data details
-
-# Model details {#sec-model-details}
-
-## Posterior predictive check
-
-In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...
-
-In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...
-
-\newpage
-
 # References

### 2024-04-24 paper/paper.qmd
Update paper
Update discussion, abstract, and introduction sections

@@
 This paper examines the implications of this rule change, comparing player statistics between the 2022-23 and 2021-22 seasons.
-We find significant correlations between shooting efficiency and MVP points, highlighting the rule's influence on player recognition.
+We found no statistically significant correlations between shooting efficiency and MVP points.
 This paper offers insights into the interplay between rule modifications, player performance, and individual accolades, underscoring the broader significance of adapting to regulatory shifts in professional sports."
@@
 To address this gap, we used the R programming language [@citeR] to analyze shooting performance metrics, such as field goal percentage (FG%), three-point percentage (3P%), free throw percentage (FT%), as well as MVP voting points won from both seasons.
-Our findings reveal patterns in player shooting performance, MVP points accumulation, and the intertwined effects of the transitional foul rule… We discovered significant correlations between shooting efficiency and MVP points, with distinct variations observed between the two seasons.
+Our findings reveal no statistically significant relationships between shooting efficiency and MVP points in the 2022-23 NBA season.
 
-Understanding the implications of the transitional foul rule in comparison to the previous season is crucial for stakeholders across the NBA landscape, from players and coaches to researchers and fans.
-By illuminating the intricate relationship between rule changes, player performance, and individual accolades, this paper contributes to a deeper understanding of the evolving dynamics within the league.
-
-This paper is structured into three main sections: Data, Results, and Discussion.
-@sec-data Data explores sources and methodology.
+This paper is structured into four main sections: Data, Model, Results, and Discussion.
+@sec-data Data explores sources, methodology, measurement, and data visualization.
 It includes various sources used for data collection, cleaning, and analyzing processes, drawing from the dataset.
-In @sec-results Results, we present our findings, focusing on data visualizations and analyses.
-@sec-discussion Discussion provides a comprehensive analysis of our research, highlights the implications for NBA basketball, the importance of limitations and considerations in our analysis, and identifies areas for future research.
-@sec-conclusion Conclusion summarizes our main findings and the broader significance of our research.
+@sec-model Model describes the statistical model employed in our analysis.
+In @sec-results Results, we present our findings, focusing on comprehensive analyses obtained from the model.
+@sec-discussion Discussion provides a summary of our research, highlights the implications for NBA basketball, the importance of limitations and considerations in our analysis, and identifies areas for future research.
 
@@
 The data analysis involved using the R programming language [@citeR], along with the `tidyverse` [@citeTidyverse], `janitor` [@citeJanitor], `knitr` [@citeKnitr], `cowplot` [@cowplot], `broom` [@broom], and `rstanarm` [@rstanarm] packages for data collection, cleaning, analysis, and statistical modeling.
-Within the `tidyverse` package [@citeTidyverse], I used `readr` [@citeReadr], `dplyr` [@citeDplyr], `tidyr` [@citeTidyr], and `ggplot2` [@citeGgplot2] for reading CSV files, manipulating data, tidying data, and creating figures.
+Within the `tidyverse` package [@citeTidyverse], `readr` [@citeReadr], `dplyr` [@citeDplyr], `tidyr` [@citeTidyr], and `ggplot2` [@citeGgplot2] are used for reading CSV files, manipulating data, tidying data, and creating figures.
 
@@
 ## Data Visualization
+
 ### The distribution of MVP points won by top 10 players in the 2021-22 and 2022-23 seasons
@@
 # Results {#sec-results}
+
 ## Summary of regression model results
+
 Our results are summarized in @tbl-coefficients and @tbl-additional.
@@
 ## Observed versus predicted MVP points won
+
 @fig-modelresults shows the observed MVP points (actual outcomes) against predicted MVP points (outcomes estimated by the model).
@@
 Conversely, an observed value of 3 has an inaccurate predicted value of 369.98.
-This means there is a potential issue of reliability and precision in the model’s predictions for top-performing players.
+This indicates that the observed MVP points and the predicted MVP points are not well aligned.
+Specifically, the model's predictions deviate significantly from the actual outcomes.
 
@@
 
-## First discussion point {#sec-first-point}
+## Summary of key findings {#sec-first-point}
 
-If my paper were 10 pages, then should be be at least 2.5 pages. The discussion is a chance to show off what you know and what you learnt from all this.
+The primary objective of this paper is to analyze the consequences of the 2022-23 NBA transitional foul rule on player shooting performance and MVP points won.
+The methodology involves collecting and analyzing data from the NBA seasons of 2021-22 and 2022-23, sourced from @basketball.
+We compare player shooting performance across these two seasons to visualize the data, illustrating trends between the seasons.
+The key variables analyzed include FG%, 3P%, FT%, and MVP points.
+There's a shift in MVP points distribution, with top-ranked players in the 2022-23 season generally receiving more points compared to the previous season, suggesting potential changes in voter perception or player performance (see @fig-points).
 
-## Second discussion point
+To further examine the relationship between shooting performance metrics and MVP voting outcomes, a linear regression model is constructed.
+This model uses MVP points as the dependent variable and shooting percentages (FG%, 3P%, FT%) as independent variables.
+The use of linear regression allows for quantifying the potential impact of shooting performance on MVP voting, considering the influence of the transitional foul rule.
+However, the results show that there are no significant findings about relationships between shooting percentages and MVP points (see @tbl-coefficients and @tbl-additional).
 
-## Third discussion point
+## The psychological impacts of the new NBA's transitional foul rule change on players
+
+To understand the implications of the NBA's transitional foul rule change, it is essential to consider its psychological impacts on players, akin to the effects observed with strategic timeouts.
+As discussed by @timeOut, strategic timeouts are often employed to disrupt a player's performance during critical moments of a game.
+While they are intended to disrupt a player's rhythm and impair performance, they often fail to achieve their intended effect [@timeOut].
+Instead, players may adapt and focus more intensely.
+Similarly, with the new transitional foul rule in the NBA, players might initially experience disruptions in their game strategy but adapt quickly to optimize performance under the new conditions.
+This adaptability is evident in the lack of significant changes in the Three-Point Percentage (3P%) and Free Throw Percentage (FT%) among the top players (@fig-percentage).
+
+Furthermore, the transitional foul rule could increase players' confidence during fast-break situations because they know that unnecessary contact could result in significant penalties for the opposing team.
+This may lead to a more fluid approach to offensive transitions, enhancing both individual and team scoring efficiency.
+It also allows players to play the game more freely and focus more on their performance rather than feeling defensive pressure.
+This trend is indicated by @fig-percentage, which shows an observable increase in the FG% of the top 10 players in the 2022-23 season compared to the previous season.
+The transitional foul rule in the NBA not only modifies the game's structural dynamics but also has profound psychological effects on players, influencing how players strategize and execute during games.
+As they adapt to these new changes, the overall impact on performance metrics and psychological resilience becomes a fascinating aspect of professional basketball dynamics.
+
+## Potential biases in the MVP voting process
+
+The inherent biases could affect the MVP voting process when evaluating the implications of the NBA's transitional foul rule changes.
+Although shooting percentages statistical metrics provide quantitative measures of player performance, there is no significant relationship between a player’s shooting performance and MVP points won, as shown by @fig-modelresults.
+Therefore, this means that other factors influence the allocation of MVP points, which may involve subjective judgment.
+
+The MVP voting process is conducted by a total of 100 sportswriters and broadcasters chosen by the NBA from the United States and Canada [@mvpVoting].
+Although these panelists are responsible for assessing player performances over the season [@mvpVoting], their subjective opinions and personal biases could affect their voting choices.
+@biases highlights similar biases in referee decisions, providing a parallel to potential biases in MVP voting.
+The research suggests that well-known NBA players often receive favorable calls, particularly during crucial moments in playoff games [@biases].
+This observation about referee bias can be extended to the MVP voting process.
+As referees may show favoritism towards star players, MVP voters can also be influenced by a player's fame, media visibility, or the prevailing narrative about their season performance.
+This bias can result in equally deserving but less prominent players not receiving comparable recognition in the MVP voting process.
+The MVP voting process not only reflects statistical performance but is also shaped by the subjective interpretations and biases of the voting panel.
 
@@
 
-Weaknesses and next steps should also be included.
+Although our analysis aims to explore the relationships between the NBA rule change, shooting performance metrics, and MVP points, it is important to acknowledge potential weaknesses that arise.
+These weaknesses include the limited scope of variables considered, the size of the datasets, statistical significance, and the subjectivity in MVP voting.
+In our findings, we did not observe a statistically significant relationship between MVP points and a player's shooting performance.
+This suggests that factors beyond shooting efficiency, such as scores, defensive contributions, team success, and player injuries, also influence MVP considerations.
+The exclusion of these variables may limit the depth and comprehensiveness of the analysis.
+In addition, our analysis is based on data from the 2022-23 and 2021-22 NBA seasons, which may not provide a sufficiently large sample size to draw definitive conclusions.
+Therefore, expanding the time frame to include data from a longer period, such as 10 years, could enhance the robustness of our key findings.
+This can allow us to comprehensively understand the impact of rule changes on player performance and MVP points.
+Moreover, the MVP voting process is subjective and may be influenced by the personal biases of the voting panel.
+This subjectivity leads to uncertainty in the interpretation of MVP points outcomes.
 
-\newpage
-
-# Conclusion {#sec-conclusion}
+To deepen our understanding of the factors influencing MVP points and the effects of rule changes in the NBA, future research is needed.
+Firstly, a broader range of player performance metrics beyond shooting percentages is necessary to grasp the player's overall impact on the game.
+Secondly, understanding the correlation between team success and individual player performance in MVP considerations is essential, as MVP voting considers a player's contribution to their team's success.
+Thirdly, extending the time frame of datasets to identify long-term trends in MVP voting and player performance is necessary to track the evolution of player recognition over time.
+Finally, exploring the composition and dynamics of the MVP voting panel, including individual biases, preferences, and voting patterns, will facilitate a deeper understanding of their influence on MVP outcomes.
+This investigation can encompass demographic factors, geographic biases, and media narratives that may impact voting decisions.
 

### 2024-04-25 paper/paper.qmd
Update introduction section
Update key findings

@@
 To address this gap, we used the R programming language [@citeR] to analyze shooting performance metrics, such as field goal percentage (FG%), three-point percentage (3P%), free throw percentage (FT%), as well as MVP voting points won from both seasons.
-Our findings reveal no statistically significant relationships between shooting efficiency and MVP points in the 2022-23 NBA season.
+
+Our results in FG% showed a positive but statistically insignificant relationship with MVP points.
+3P% had a negative relationship with MVP points, also statistically insignificant.
+FT% showed a positive but statistically insignificant relationship with MVP points.
+The model had a high residual standard error (314.30), indicating substantial unexplained variability, and a low Adjusted R-squared (0.17), suggesting potential overfitting or irrelevance of some variables.
+Therefore, our findings reveal no statistically significant relationships between shooting efficiency and MVP points in the 2022-23 NBA season.
 
